---
title: "Muni Code Correction"
author: "Michelle Evans"
date: "8/2/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rgdal)
library(dplyr)
```

This document provides the reasoning for any municipal code corrections in the YF data. The shapefile provided by the IBG (Brazilian govt) did not include some of the codes that were in the YF case data, so we needed to figure out a way to merge them.

#Identifying Problem Municipalities

```{r load shapefile}
brazShape <- readOGR("../data_clean", "BRAZpolygons") #this has appropriate codes
```

```{r load YF data}
yfData <- readRDS("../data_clean/YFcases/YFlong.rds")
```

```{r}
yfMuni <- unique(yfData$muni.no)
shapeMuni <- as.character(unique(brazShape@data$muni_no))

missingMuni <- unique(yfData[!(yfMuni %in% shapeMuni), 3:4]) #33 total (this includes the nearly unknown and imported)

#drop unknowns (codes for state unknowns end with 4 zeros)
#some with 5 zeros aren't caught, and neither is one that is just 0
missingMuni <- missingMuni[!((substr(missingMuni$muni.no, 3,6) == "0000")| (substr(missingMuni$muni.no, 3,6) == "+05") | missingMuni$muni.no==0),] #results in 11 total

#several of these are just misspellings becuase of accents
unique(missingMuni$muni.no) 

#drop duplicates
missingMuni <- missingMuni %>%
  group_by(muni.no) %>%
  slice(1) %>%
  ungroup() #new total is only 6 muni's

#check if these missingMuni's have any cases
sum(yfData$case[yfData$muni.no %in% missingMuni$muni.no], na.rm=T) #no cases
```

# One by One Explanation and Reasoning

```{r}
missingMuni$old.no <- NA
```


```{r}
missingMuni[1,]
missingMuni$old.no[1] <- 150680
```

Mojui dos Campos is in Para and became a municipality in 2013. It was originally part of Santarem (150680).


```{r}
missingMuni[2,]
missingMuni$old.no[2] <- 221100
```

Nazaria is in Piaui, Teresina. It was created from Teresina (221100) in 2013.

```{r}
missingMuni[3,]
missingMuni$old.no[3] <- 420460
```

Balneario Rincao is in Santa Catarina. It became a municipality in 2013. It was in the sub-municipality of Icara, which was in Criciuma (420460).

```{r}
missingMuni[4,]
missingMuni$old.no[4] <- 420940
```

Pexcaria Brava is also in Santa Catarina. It became a municipality in 2009 from Laguna (420940).

```{r}
missingMuni[5,]
missingMuni$old.no[5] <- 430210
```

Pinto Bandeira is Rio Grande do Sul. It became a municipality in 2001, but then in 2003 lost its municipality status. It then became a municipality again in 2013, from Bento Goncalves (430210). The IBGE website claims its territory rested within Pinto Bandeira until 2013, so I don't think we need to worry about it in those early years.

```{r}
missingMuni[6,]
missingMuni$old.no[6] <- 500325
```

Paraiso das Aguas is in Mato Grosso do Sul. It became a municipality in 2003, from Agua Clara, Costa Rica, and Chapadao do Sul. However,  there was a suit from Agua Clara, so it didn't become independent until 2013. Technically it was a district of Costa Rica (500325), so I think it is safe to say it came from that.