---
title: "Muni Code Correction"
author: "Michelle Evans"
date: "8/2/2017"
output: html_document
---

```{r setup, include=FALSE, eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rgdal)
library(dplyr)
```

This document provides the reasoning for any municipal code corrections in the YF data. The shapefile provided by the IBG (Brazilian govt) did not include some of the codes that were in the YF case data, so we needed to figure out a way to merge them.

#Identifying Problem Municipalities

```{r load shapefile}
brazShape <- readOGR("../data_clean", "BRAZpolygons") #this has appropriate codes
```

```{r load YF data}
yfData <- readRDS("../data_clean/YFcases/YFlong.rds")
```

```{r}
yfMuni <- unique(yfData$muni.no)
shapeMuni <- as.character(unique(brazShape@data$muni_no))

missingMuni <- unique(yfData[!(yfMuni %in% shapeMuni), 3:4]) #36 total (this includes the nearly unknown and imported)

#drop unknowns (codes for state unknowns end with 4 zeros)
#some with 5 zeros aren't caught, and neither is one that is just 0
missingMuni <- missingMuni[!((substr(missingMuni$muni.no, 3,6) == "0000")| (substr(missingMuni$muni.no, 3,6) == "+05") | missingMuni$muni.no==0),] #results in 9 total

#several of these are just misspellings becuase of accents
unique(missingMuni$muni.no) 

#drop duplicates
missingMuni <- missingMuni %>%
  group_by(muni.no) %>%
  slice(1) %>%
  ungroup() #new total is only 6 muni's

#check if these missingMuni's have any cases
sum(yfData$case[yfData$muni.no %in% missingMuni$muni.no], na.rm=T) #no cases
```

# One by One Explanation and Reasoning

```{r}
missingMuni$mother <- NA
missingMuni$emancipationYear <- NA

```


```{r}
missingMuni[1,]
missingMuni$mother[1] <- 150680
missingMuni$emancipationYear[1] <- 2013
```

Mojui dos Campos is in Para and became a municipality in 2013. It was originally part of Santarem (150680).


```{r}
missingMuni[2,]
missingMuni$mother[2] <- 221100
missingMuni$emancipationYear[2] <- 2013
```

Nazaria is in Piaui, Teresina. It was created from Teresina (221100) in 2013.

```{r}
missingMuni[3,]
missingMuni$mother[3] <- 420460
missingMuni$emancipationYear[3] <- 2013
```

Balneario Rincao is in Santa Catarina. It became a municipality in 2013. It was in the sub-municipality of Icara, which was in Criciuma (420460).

```{r}
missingMuni[4,]
missingMuni$mother[4] <- 420940
missingMuni$emancipationYear[4] <- 2009
```

Pexcaria Brava is also in Santa Catarina. It became a municipality in 2009 from Laguna (420940).

```{r}
missingMuni[5,]
missingMuni$mother[5] <- 430210
missingMuni$emancipationYear[5] <- 2013
```

Pinto Bandeira is Rio Grande do Sul. It became a municipality in 2001, but then in 2003 lost its municipality status. It then became a municipality again in 2013, from Bento Goncalves (430210). The IBGE website claims its territory rested within Pinto Bandeira until 2013, so I don't think we need to worry about it in those early years.

```{r}
missingMuni[6,]
missingMuni$mother[6] <- 500325
missingMuni$emancipationYear[6] <- 2013
```

Paraiso das Aguas is in Mato Grosso do Sul. It became a municipality in 2003, from Agua Clara, Costa Rica, and Chapadao do Sul. However,  there was a suit from Agua Clara, so it didn't become independent until 2013. Technically it was a district of Costa Rica (500325), so I think it is safe to say it came from that.

```{r}
missingMuni[7,]
missingMuni$mother[7] <- 500295
missingMuni$emancipationYear[7] <- 2009 
```

Paraiso das Aguas was emancipated from Chapadao do Sul (500295) in 2009. It was potentially annexed from 3 different municipalities (Água Clara, Costa Rica and Chapadão do Sul). The mother muni was based on a judgement call using a map. 

```{r}
missingMuni[8,]
missingMuni$mother[8] <- 510800
missingMuni$emancipationYear[8] <- 2000 
```

Ipiranga do Norte was emancipated from Tapurah (510800) in 2000.


```{r}
missingMuni[9,]
missingMuni$mother[9] <- 510800
missingMuni$emancipationYear[9] <- 2000 
```

Itanhanga was emancipated from Tapurah (510800) in 2000

##Additional muni
Pinto Banderia (431454) was absorbed and then emancipated from Bento Goncalves (430210). The first in 2001 and then emancipated in 2012.  

Save our muni correction file to use in the model markdown file.

```{r}
write.csv(missingMuni, "../data_raw/demographic/muniCorrections.csv", row.names = F)
```

