---
title: "Normalizing and Building Dataset"
author: "Reni Kaul"
date: "5/31/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Building the data set

This file builds the data set used for the YF model. The variables are currently saved seperately with entries identified by muni.no and month.no. Scaled variables are also created in the process. The normality of variable distributions are visually checked. 

- case :reporting of any YF cases (0,1)
- numCase :number of reported YF cases 

- ndvi : NDVI for that month
- ndviScale : NDVI rescaled to max value for that muni and calendar month

- popLog10 : population density; log10 transformed

- rf : mean hourly rainfall; cube root transformed
- rfScale : mean hourly rainfall rescaled to max value for that muni and calendar month; cube root transformed

- temp : mean monthly air temperature 
- tempScale : mean monthly air temperature rescaled to max value for that muni and calendar month

- fire : density of fires oberserved in month per muni; cube root transformed
- fireScale : density of fires oberserved in month per muni rescaled to max value for that muni and calendar month. NA values converted to zero; cube root transformed 

- spRich : number of non-human primates by species with ranges based on IUCN {0-22}; cube root transformed 

- primProp : sum of each municipalities relative area that is both agricultural and falls within a primate genus range. {0,9}; cube root transformed

- vectorOcc : maximum probability of a known spillover vector occuring within that municipality {0,1}

- muni.no : unique number to identify municipality
- month.no : numbered month of observation {1,168}
- muni.name : character string name
- cal.month : calendar month {1,12}
- year : year {2001,2014}


```{r library and functions}
library(dplyr)
library(gplots)
library(ROCR)
library(pROC)
library(rgdal)
library(rgeos)
library(BalancedSampling)
library(scatterplot3d)
library(rcompanion) # useful for plotNormalHistogram fnc

## function to explore transformation
ExploreTransformations = function(data, VarName = "", shift = FALSE){
  if(shift == TRUE){
  transLog <- log(data + 1)
  } else {
    transLog <- log(data)
  }
  transSrt <- sqrt(data)
  transCube <- sign(data) * abs(data)^(1/3)
  par(mfrow=c(2,2))
  plotNormalHistogram(data, main= "original")
  plotNormalHistogram(transLog, main = "Log")
  plotNormalHistogram(transSrt, main = "Srt")
  plotNormalHistogram(transCube, main = "Cube")
  title(VarName, outer=TRUE, line= -1)
}

TranCube = function(x) {sign(x) * abs(x)^(1/3)}

```

```{r load dynamic data}
# Dynamic values for muni's (trimmed to study range)
pop <- readRDS("../data_clean/demographic/pop.rds") %>% filter(year<2014)
rf <- readRDS("../data_clean/environmental/allRainfall.rds") %>% filter(year<2014)
fires <- readRDS("../data_clean/environmental/numFires.rds") %>% filter(year<2014)
ndvi <- readRDS("../data_clean/environmental/allNDVI.rds") %>% filter(year<2014)
temp <- readRDS("../data_clean/environmental/meanTemperature.rds") %>% filter(year<2014)
```


```{r load static like data}

# Static value for muni's 
NHPProp <- readRDS("../data_clean/environmental/primProp.rds") #changes every year, missing 2014
NHPRichness <- readRDS("../data_clean/environmental/primRichness.rds") #doesn't change with time
mosi.occurence <- readRDS("../data_clean/environmental/mosiProb.rds")

```

# Check the distribution of covariates
Check different transformations of covariates against a normal distribution. The natural log, square root, and cube root transformations were applied and then visually compared to a normal distribution with the same mean (blue line in plots). In cases where the original data contained zeros the log transformation was `log(x+1)`

**Transformation summary** 

Unaltered variables:

- NDVI*
- Temperature*
- Mosi occurance

Cube root transformed:

- Rainfall*
- Fire density*
- NHP richness
- NHP/Ag Overlap

Log10 transformed:

- Population density

Details below. 

### Population density 

Population density is by year. and should be $log_{10}$ transformed. 

```{r population trans}
par(mfrow=c(1,2))
  plotNormalHistogram(pop$densitypop, main= "original")
  plotNormalHistogram(log10(pop$densitypop), main = "Log10")
  title("Population Density", outer=TRUE,  line = -1)

```

### Rainfall

Mean hourly rainfall and scaled mean hourly rainfall should both be cube root transformed. 
```{r rainfall trans}
rf <- rf %>% 
  group_by(muni.no, cal.month) %>%
  mutate(RFScale = hourlyRainfallMean/max(hourlyRainfallMean))

ExploreTransformations(rf$hourlyRainfallMean, VarName = "Rainfall", shift = TRUE)

ExploreTransformations(rf$RFScale,VarName = "Rainfall Scale", shift = TRUE)

```

### Fire density 

Fire density is ugly despite any transformation. For consistency, we should cube root transform. 
```{r fire trans}
fires <- fires %>%
      group_by(muni.no, cal.month) %>%
   mutate(fireDenScale = ifelse(max(fireDens)==0,0,fireDens/max(fireDens))) %>% #scale fire densities to muni and month
   ungroup() 

ExploreTransformations(fires$fireDens,  VarName = "Fire Density", shift = TRUE)
ExploreTransformations(fires$fireDenScale, shift = TRUE)

```

### NDVI

Tranformations aren't needed for NDVI or NDVI scaled values. 
```{r NDVI trans}
ndvi <- ndvi %>%
  group_by(muni.no, cal.month) %>%
  mutate(NDVIScale = NDVI/max(NDVI)) %>% #scale NDVI to muni and month
  ungroup()

ExploreTransformations(ndvi$NDVI,  VarName = "NDVI")
ExploreTransformations(ndvi$NDVIScale, VarName = "NDVI Scale")
```

### Temp 

Transformations aren't needed for temp or temp scale
```{r temp trans}
temp <- temp %>%
  group_by(muni.no, cal.month) %>%
  mutate(tempScale = tempMean/max(tempMean)) %>%
  ungroup()

ExploreTransformations(temp$tempMean, VarName = "Temp")
ExploreTransformations(temp$tempScale, VarName = "Temp Scale")

```

### NHP Prop

Primate proportion should be cube root transformed

```{r NHP trans}
ExploreTransformations(NHPProp$primProp, shift = TRUE)
```

### Species Richness

Species richness should be cube root transformed. 
```{r}
ExploreTransformations(NHPRichness$spRich, shift = TRUE)
```

### Mosi Occurence 

Mosi occurance does not need to be transformed. 

```{r}
ExploreTransformations(mosi.occurence$vector.prob, shift =TRUE)
```



# Assemble the data

```{r cleaning environmental data, eval = FALSE}
all.enviro <- plyr::join_all(list(ndvi, pop, rf, temp, fires), 
                           by=c("muni.no", "month.no"), type="full") %>%
              select(-c(muni, muniArea, area, pop, numFire))

environCovar <- all.enviro %>%
  # scaled variables 
    group_by(muni.no, cal.month) %>%
    mutate(tempScale = tempMean/max(tempMean)) %>%
    mutate(ndviScale = NDVI/max(NDVI)) %>%
    mutate(rfScale = hourlyRainfallMean/max(hourlyRainfallMean)) %>%
    mutate(fireScale = ifelse(max(fireDens)==0,0,fireDens/max(fireDens))) %>%
    ungroup() %>%
  # transformations with name changes
    mutate(popLog10 = log10(densitypop)) %>%
    mutate(rf = TranCube(hourlyRainfallMean)) %>%
    mutate(rfScale = TranCube(rfScale)) %>%
    mutate(fire = TranCube(fireDens)) %>%
    mutate(fireScale = TranCube(fireScale)) %>%
  # just name changes
    mutate(ndvi = NDVI) %>%
    mutate(temp = tempMean) %>%
  # drop the extra things
    select(c(muni.no, muni.name, year, cal.month, month.no, rf, rfScale, ndvi, ndviScale, temp, tempScale, popLog10, fire, fireScale))
                
```

```{r cleaning static data, eval = FALSE}
NHPenvironCovar <- NHPRichness %>%
    mutate(spRich = TranCube(spRich)) %>%
    select(-c(muni.name)) %>%
    dplyr::right_join(environCovar, by=c('muni.no')) 

NHPenvironCovar <-  NHPProp %>%
    mutate(primProp = TranCube(primProp)) %>%
    select(-c(muni.name)) %>%
    dplyr::right_join(NHPenvironCovar, by=c('muni.no','year')) 

allCovar <- mosi.occurence %>%
    mutate(vectorOcc = vector.prob/1000) %>%
    select(-c(vector.prob)) %>%
    dplyr::right_join(NHPenvironCovar, by=c('muni.no')) 

```


## YF case data

```{r YF case, eval = FALSE}
## YF cases------------------------------ 
cases <- readRDS("../data_clean/YFcases/YFlong.rds")
 #dimensions: Entries for all muni ONLY if a case reported in any muni that month
cases[is.na(cases)] <- 0 #replace NA with zero  

cases <- cases %>%
  filter(year<2014) %>%
  #drop annual totals
  filter(cal.month<=12) %>%
  #drop unknown origin totals %>%
  filter(!grepl("ignorado", muni)) %>%
  mutate(numCase=case) %>% 
  mutate(case=ifelse(numCase==0,0,1))


#add in cases
allData <- cases %>%
            select(c(muni.no, month.no, case, numCase)) %>%
            dplyr::right_join(allCovar, by=c('muni.no', 'month.no')) %>%
            mutate(case=ifelse(is.na(case),0,case)) %>% #replace NA with zero
            mutate(numCase=ifelse(is.na(numCase) == TRUE,0,numCase)) #replace NA with zero
   
saveRDS(allData, "../data_clean/FinalData_Mosi/allData.rds") 
# rm(list = ls())
allData <- readRDS("../data_clean/FinalData_Mosi/allData.rds")
```

# Spliting data

Data splits made based on old dataset muni.no and month.no

```{r , eval = FALSE}
# Pull in the old data

oldTest <- readRDS("../data_clean/TestingDataSpat2.rds") 
oldTestLow <- readRDS("../data_clean/TestingDataLowNHP2.rds") 
oldTestHigh <- readRDS("../data_clean/TestingDataHighNHP2.rds")

# pull out rows in common between old data set and new
testing <-  allData %>% semi_join(oldTest, by=c("muni.no", "month.no"))
testingLow <- allData %>% semi_join(oldTestLow, by = c("muni.no", "month.no"))
testingHigh <- allData %>% semi_join(oldTestHigh, by = c("muni.no", "month.no"))

# save testing data
saveRDS(testing, file = "../data_clean/FinalData_Mosi/TestingData.rds")
saveRDS(testingLow, "../data_clean/FinalData_Mosi/TestingDataLowNHP.rds")
saveRDS(testingHigh, "../data_clean/FinalData_Mosi/TestingDataHighNHP.rds")

# Repeat steps for training 

# pull in the old data
oldTrain <- readRDS("../data_clean/TrainingDataSpat2.rds")
oldTrainLow <- readRDS("../data_clean/TrainingDataLowNHP2.rds")
oldTrainHigh <- readRDS("../data_clean/TrainingDataHighNHP2.rds")

# pull out rows in common between old data set and new
training <-  allData %>% semi_join(oldTrain, by=c("muni.no", "month.no"))
trainingLow <- allData %>% semi_join(oldTrainLow, by = c("muni.no", "month.no"))
trainingHigh <- allData %>% semi_join(oldTrainHigh, by = c("muni.no", "month.no"))

# Save data with spatial and temporal stratification
#individual training and testing data
saveRDS(training, file="../data_clean/FinalData_Mosi/TrainingData.rds")
saveRDS(trainingLow, "../data_clean/FinalData_Mosi/TrainingDataLowNHP.rds")
saveRDS(trainingHigh, "../data_clean/FinalData_Mosi/TrainingDataHighNHP.rds")

```


```{r sum of cases, eval = FALSE}
CaseSummary = function(x){
  pos <- sum(x$case)
  total <- dim(x)[1]
  bg <- total - pos
  return(c(pos,bg,total))
}

# uick sanity check. 
CaseSummary(training)
CaseSummary(trainingLow)
CaseSummary(trainingHigh)


CaseSummary(testing)
CaseSummary(testingLow)
CaseSummary(testingHigh)
```

