---
title: "Covar Exploration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rgdal)
library(raster)
library(rgeos)
library(corrplot)
library(GGally)
```


```{r load data}
brazil <- readOGR("../data_clean", "BRAZpolygons")

covariates <- readRDS("../data_clean/FinalData.rds")
```

# Basic Visualizations

## Correlation Amongst Covariates

We wanted to explore the correlation between covariates, particularly as many are features created from each other.

The primate proportion data doesn't have 2014, generating NAs. Becuase this hasn't come up as an informative variable in our model so far, I'm just ignoring it for now.

```{r}
corrplot(cor(covariates[, 3:15]), method = "number")
```

All of the rainfall variables are extremely correlated (>0.95), suggesting we should only use one of those in the model. The rest seem okay.

There is also a way to scatterplot them all against each other. You cannot do all of them at once (because it is millions and millions of points), so I have chosen some to explore based on the plot above. In this case, they are colored by if there was a case or not. The device driver also can't deal with it, so it is better to save to a seperate image. It will take 2-3 minutes per variable (increasing exponentially too), and I wouldn't include more than 8 unless on a high-mem machine.

```{r}
#reorder so that lines with cases are on top
covariates <- covariates[order(covariates$case, decreasing = F),]
#drop NAs in case data
covariates <- dplyr::filter(covariates, !is.na(case))
covariates$caseSort <- as.factor(covariates$case) #variable for color
```


```{r}
png("pairwisePlotBase.png", res = 400, width = 8, height = 8, units ="in")
graphics::pairs(covariates[,c('NDVIScale', 'popLog10', "spRich", "RFScale", "fireDenSqrt", "tempScale")], 
                col = c("black", "red")[covariates$case+1], pch=20)
dev.off()
```

We can also do it with 'ggpairs' to get an idea of the density of data at the same time (since some trends that stand out may be irrelevant when considering distribution of data). Just note that ggpairs really sucks sometimes as far as configuration, and is slightly slower than base graphics.

```{r}
png("pairwisePlotPairs.png", res = 400, width = 8, height = 8, units ="in")
p <- ggpairs(covariates, columns = c('NDVIScale', 'popLog10', "spRich", "RFScale", "fireDenSqrt", "tempScale"), 
        diag = list(continuous = "density"), 
        mapping = aes(color = caseSort, alpha = 0.5))

#hack colors
for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] + 
        scale_fill_manual(values=c("black", "red")) +
        scale_color_manual(values=c("black", "red"))  
  }
}

p

dev.off()
```


## Covariates Over Time

# Temporal Visualization

The following function creates the anomaly images similar to those from Lowe et al. (2017), with the choice of choosing a municipality, state, or whole nation to view over.

```{r}
anomalyViz <- function(agg, id, vars){
  #' Visualize Temporal Anomalies
  #'
  #' This function returns plots of temporal anomalies over different aggregations and variables
  #'
  #' @param agg What level of aggregation do you want? "muni", "state", "nation"
  #' @id only if agg is muni or state, the identifying codes for which one you want to visualize
  #' @vars which variables you want to visualize, based on the column names of the covariate dataframe
  
  require(dplyr)
  require(viridis)
  require(RColorBrewer)
  # subset out data to meet those defined in function arguments
  if (agg=="muni"){
    data2plot <- covariates %>%
      filter(muni.no==id) %>%
      select(muni.no, "year", cal.month, vars)
    
    cases <- covariates %>%
      filter(muni.no==id) %>%
      select(muni.no, "year", cal.month, case)
    
  } else if (agg=="state"){
    
    id <- as.numeric(substr(id, 1,2)) #get state id from muni.no just in case
    
    data2plot <- covariates %>%
      mutate(state = as.numeric(substr(muni.no,1,2))) %>%
      filter(state == id) %>%
      select(state, "year", cal.month, vars) %>%
      group_by(state, year, cal.month) %>%
      summarise_all(mean, na.rm=T) %>%
      ungroup()
    
      cases <- covariates %>%
        mutate(state = as.numeric(substr(muni.no,1,2))) %>%
        filter(state == id) %>%
        select(state, "year", cal.month, case) %>%
        group_by(state, year, cal.month) %>%
        summarise_all(sum, na.rm=T)
               
  } else if (agg=="nation"){
    
    id  <- NULL
    
    data2plot <- covariates %>%
      select("year", cal.month, vars) %>%
      group_by(year, cal.month) %>%
      summarise_all(mean, na.rm=T) %>%
      ungroup()
    
    cases <- covariates %>%
      select("year", cal.month, case) %>%
      group_by(year, cal.month) %>%
      summarise_all(sum, na.rm=T)
  }
  #adjust weird NAs in cases
  cases$case[is.na(cases$case)] <- 0
  
  for (i in 1:length(vars)){
    varPlot <- vars[i]
    x <- 1:12
    y <- unique(data2plot$year)
    dataVec <- as.numeric(unlist(data2plot[,varPlot])) #fix weird list thing
    z <- array(dataVec, c(length(x), length(y)))
    filled.contour(x,y,z, 
                 col = rev(brewer.pal(7,'RdBu')),
                 plot.title = title(main = paste(varPlot, "in", agg, id)),
                 xlab = "Month", 
                 ylab = "Year",
                 plot.axes = {axis(1,at=c(1:12),labels=c("J","F","M","A","M","J","J","A","S","O","N","D"));
                              axis(2,min(data2plot$year):max(data2plot$year));
                              points(x=cases$cal.month, y=cases$year, cex=cases$case/5)
                   },
                 nlevels = 7
                 )
    #mtext(paste(varPlot, "in", agg, id))
  }
  
} #close function
```

Test it out. Sometimes need to run dev.off to get labels to work appropriately. Circles represent cases at that aggregation scale (summed).
```{r}
anomalyViz(agg="state", id=521690, vars = c("RFScale", "fireDens", "tempScale"))
```

# Spatial Visualization


