---
title: "Manuscript Figures"
author: ""
date: "January 23, 2018"
output: html_document
---

# Packages and Set-Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)



#fig 3
library(sf)
library(raster)
library(ggplot2)

#fig 4
library(sf)
library(raster)
library(viridis)
library(cowplot)
library(ggthemes)
library(lubridate)
```

This document creates the figures for:

*Insert manuscript citation here*

I'm currently loading data individually for each figure, then if you want to remake one figure you don't have to rerun the whole document, just that section.

# Figure 1: Variable Importance

Dodged barplot of variable importance

# Figure 2: Coord Plot

"weird ladder"


#Figure 3: Maps of Calendar Months

This map is a 3 x 2 lattice, showing three characteristic annual time points in the year (represented as mean across all instances of that month) across the single and dual-process model. These values are re-scaled in order to be comparable across models. Several ways we have done this: probability x 10^3 (multiply raw output by ratio of positive/background in the model, then multiply by 10^3 so it is easier to see), normalized (scaled to maximum value output by the model), and frequency x 10^3 (rescaled by constant c, so that the sum of all of the predictions multipled by c equals the initial number of positives).

## Load Data

```{r}
braz.map <- st_read("../data_clean", "BRAZpolygons")
braz.map <- rename(braz.map, muni.no = muni_no)

# NHP split index
nhp.split <- readRDS("../data_clean/environmental/twoModelSplit.rds")
nhp.split.map <- left_join(braz.map, nhp.split, by = "muni.no") %>%
  group_by(above5split) %>%
  summarize() %>%
  st_buffer(0.01)

#get brazil outline
braz.outline <- getData('GADM', country ='BRA', level = 0)
braz.outline <- dplyr::select(st_as_sf(braz.outline), ISO)

# prediction data
pred.one.model <- readRDS("../data_out/MS_results/OneModel/wholePredictions.rds")
pred.highNHP <- readRDS("../data_out/MS_results/HighModel/wholePredictions.rds")
pred.lowNHP <- readRDS("../data_out/MS_results/LowModel/wholePredictions.rds")
```

## Calculate Data for Plots

Rescale low and high NHP data and combine into one dataframe.

```{r}
pred.lowNHP <- pred.lowNHP %>%
  mutate(prob = prediction * (sum(case)/n())*1000) %>%
  mutate(pred.norm = prediction/max(prediction)) %>%
  mutate(pred.freq = prediction * (sum(case)/sum(prediction))*1000) %>%
  mutate(model = "two.model")

pred.highNHP <- pred.highNHP %>%
  mutate(prob = prediction * (sum(case)/n())*1000) %>%
  mutate(pred.norm = prediction/max(prediction)) %>%
  mutate(pred.freq = prediction * (sum(case)/sum(prediction))*1000) %>%
  mutate(model = "two.model")

pred.one.model <- pred.one.model %>%
  mutate(prob = prediction * (sum(case)/n())*1000) %>%
  mutate(pred.norm = prediction/max(prediction)) %>%
  mutate(pred.freq = prediction * (sum(case)/sum(prediction))*1000) %>%
  mutate(model = "one.model")

pred.all <- rbind(pred.lowNHP, pred.highNHP, pred.one.model) %>%
  mutate(cal.month = month.no %% 12)

pred.all$cal.month[pred.all$cal.month==0] <- 12

pred.all.month <- pred.all %>%
  group_by(muni.no, model, cal.month) %>%
  summarise(mean.prob = mean(prob), 
            mean.pred.norm = mean(pred.norm),
            mean.pred.freq = mean(pred.freq))


pred.map <- left_join(pred.all.month, braz.map, by = "muni.no")
  
```


## Plot

```{r static map theme}
theme_map <- function(...) {
  theme_minimal() +
  theme(
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "right",
    panel.grid = element_blank(),
    plot.background = element_rect(fill = NA, color = NA), 
    panel.background = element_rect(fill = NA, color = NA), 
    legend.background = element_rect(fill = NA, color = NA),
    strip.background = element_blank(),
    strip.text = element_blank(),
    ...
  )
}
```

```{r plot all months, eval = F}
for (k in unique(pred.map$model)){
  
  for(i in 1:12){
    
    pdf(file = paste0("figures/Fig4/", gsub("\\.", "", k), "Month", i, ".pdf"), width = 12, height = 12, family = "sans")

    p <- ggplot(data = pred.map[pred.map$cal.month == i & pred.map$model == k,])+
      geom_sf(aes(fill = mean.pred), color = NA) +
      ggtitle(paste(k, "month", i, sep = " "))
    
    print(p)

    dev.off()

  } #month loop

} #model loop
```

Subset out months we want (Feb, June, Dec.)

```{r}
pred.map.subset <- pred.map %>%
  #filter(cal.month %in% c(2,6,12)) %>%
  filter(cal.month %in% c(2)) %>%
  mutate(monthLabel = month.abb[cal.month])
```

Plot. Because of the line denoting the split, we need to plot the single and dual process model seperately, then facet together with cowplot.

Single process.

```{r}
top.row <- ggplot() +
  geom_sf(data = pred.map.subset[pred.map.subset$model=="one.model",], 
          aes(fill = mean.pred.norm), color = NA) +
  scale_fill_viridis(option = "magma", direction = -1, discrete = F,
                     na.value = "gray70", 
                     guide = guide_colorbar(
                       title = "Prediction (normalized)",
                       direction = "vertical",
                       barheight = unit(35, units = "mm"),
                       barwidth = unit(2, units = "mm"),
                       draw.ulim = F,
                       title.position = 'top',
                       title.hjust = 0.5,
                       label.hjust = 0.5,
                       title.theme = element_text(size = 8, angle = 0),
                       label.theme = element_text(size = 8, angle = 0))) +
  facet_wrap(~monthLabel, dir = "h") +
  geom_sf(data = braz.outline, fill = NA, color = "black")
```


```{r}
bottom.row <- ggplot() +
  geom_sf(data = pred.map.subset[pred.map.subset$model=="two.model",], 
          aes(fill = mean.pred.norm), color = NA) +
  scale_fill_viridis(option = "magma", direction = -1, discrete = F,
                     na.value = "gray70", 
                     guide = guide_colorbar(
                       title = "Prediction (normalized)",
                       direction = "vertical",
                       barheight = unit(35, units = "mm"),
                       barwidth = unit(2, units = "mm"),
                       draw.ulim = F,
                       title.position = 'top',
                       title.hjust = 0.5,
                       label.hjust = 0.5,
                       title.theme = element_text(size = 8, angle = 0),
                       label.theme = element_text(size = 8, angle = 0))) +
  facet_wrap(~monthLabel, dir = "h") +
  geom_sf(data = nhp.split.map, fill = NA, color = "black")
```




```{r}
pdf(file = "figures/manuscript/Figure3.pdf", width = 6.69, height = 4.75, family = "sans")

ggplot() +
  geom_sf(data = pred.map.subset, aes(fill = mean.pred), color = NA) +
  scale_fill_viridis(option = "magma", direction = -1, discrete = F,
                     na.value = "gray70", 
                     guide = guide_colorbar(
                       title = "Probability",
                       direction = "vertical",
                       barheight = unit(35, units = "mm"),
                       barwidth = unit(2, units = "mm"),
                       draw.ulim = F,
                       title.position = 'top',
                       title.hjust = 0.5,
                       label.hjust = 0.5,
                       title.theme = element_text(size = 8, angle = 0),
                       label.theme = element_text(size = 8, angle = 0))) +
  facet_wrap(~model + monthLabel) +
  geom_sf(data = nhp.split.map, fill = NA, color = "black")

dev.off()
```

Eventually the one.model will not have the split in it. I just can't figure it out yet.

Caption: Probability of yf spillover averaged over calendar month for Dec., Feb., and Jun. Result from single model in top row and dual process model in bottom row.

# Figure 4: Lattice Plot 

Figure showing results collapsed over space and time in a lattice. 

## Load Data
```{r}
# spatial data
braz.map <- st_read("../data_clean", "BRAZpolygons")
braz.map <- rename(braz.map, muni.no = muni_no)

#get brazil outline
braz.outline <- getData('GADM', country ='BRA', level = 0)
braz.outline <- st_as_sf(braz.outline)
braz.outline <- dplyr::select(braz.outline , id = OBJECTID, ID_0, ISO)

# prediction data
pred.one.model <- readRDS("../data_out/MS_results/OneModel/wholePredictions.rds")
pred.highNHP <- readRDS("../data_out/MS_results/HighModel/wholePredictions.rds")
pred.lowNHP <- readRDS("../data_out/MS_results/LowModel/wholePredictions.rds")

```


```{r download background countries}
#cshapes package?
```


## Calculate Values

### Over Space

I'm calculating the variance per municipality to get a measure of "flickering". The variance is calculated on the raw values.

```{r}
# calculate variance per municipality over whole time period
var.one.model <- as.data.frame(pred.one.model) %>%
  group_by(muni.no) %>%
  summarise(pred.var = var(prediction)) %>%
  mutate(model = "one.model")

var.highNHP <- as.data.frame(pred.highNHP) %>%
  group_by(muni.no) %>%
  summarise(pred.var = var(prediction)) %>%
  mutate(model = "highNHP")

var.lowNHP <- as.data.frame(pred.lowNHP) %>%
  group_by(muni.no) %>%
  summarise(pred.var = var(prediction)) %>%
  mutate(model = "lowNHP")

var.all.models <- rbind(var.one.model, var.highNHP, var.lowNHP) %>%
  left_join(braz.map, by = "muni.no")

rm(var.one.model, var.highNHP, var.lowNHP)

var.all.models$model <- factor(var.all.models$model, levels = c("one.model", "lowNHP", "highNHP"))
```


### Over Time

This looks at a time series, averaged over all municipalities by month. This is plotting the probabilities. Note that this kind of creates a problem because they are on such different scales. We can make the y axis of each plot free if we want to visualize the seasonality better. Or not plot probability.

```{r}
time.one.model <- as.data.frame(pred.one.model) %>%
  mutate(pred.scale = prediction * (sum(case)/n())) %>%
  group_by(month.no) %>%
  summarise(pred.mean = mean(pred.scale, na.rm = T)) %>%
  mutate(model = "one.model")

time.lowNHP <- as.data.frame(pred.lowNHP) %>%
  mutate(pred.scale = prediction * (sum(case)/n())) %>%
  group_by(month.no) %>%
  summarise(pred.mean = mean(pred.scale, na.rm = T)) %>%
  mutate(model = "lowNHP")

time.highNHP <- as.data.frame(pred.highNHP) %>%
  mutate(pred.scale = prediction * (sum(case)/n())) %>%
  group_by(month.no) %>%
  summarise(pred.mean = mean(pred.scale, na.rm = T)) %>%
  mutate(model = "highNHP")

time.all.models <- rbind(time.one.model, time.highNHP, time.lowNHP) 

#create cal.month variable (1 - 12)
time.all.models <- time.all.models %>%
  mutate(cal.month = month.no %% 12)

time.all.models$cal.month[time.all.models$cal.month==0] <- 12

#create year variable
time.all.models <- time.all.models %>%
  mutate(year = ((month.no-1) %/% 12) + 2001) %>%
  mutate(monthLabel = month(cal.month, label = T))

#order models
time.all.models$model <- factor(time.all.models$model, levels = c("one.model", "lowNHP", "highNHP"))

rm(time.one.model, time.highNHP, time.lowNHP)
```

## Plot

### Over Space

```{r map theme}
theme_map <- function(...) {
  theme_minimal() +
  theme(
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "left",
    panel.grid = element_blank(),
    plot.background = element_rect(fill = NA, color = NA), 
    panel.background = element_rect(fill = NA, color = NA), 
    legend.background = element_rect(fill = NA, color = NA),
    strip.background = element_blank(),
    strip.text = element_blank(),
    plot.margin = unit(c(0,5,0,5), unit = "pt"),
    ...
  )
}
```

## Add gray background to this

```{r facet}
top.row <- ggplot() +
    geom_sf(data = braz.outline, color = "gray25", size = 0.1) +
    geom_sf(data = var.all.models, aes(fill = pred.var), color = NA) +
    coord_sf(datum = NA)+
    scale_fill_viridis(option = "magma", direction = -1, discrete = F,
                     na.value = "gray70", 
                     guide = guide_colorbar(
                       title = "Variance",
                       direction = "vertical",
                       barheight = unit(35, units = "mm"),
                       barwidth = unit(2, units = "mm"),
                       draw.ulim = F,
                       title.position = 'top',
                       # some shifting around
                       title.hjust = 0.5,
                       label.hjust = 0.5,
                       title.theme = element_text(size = 8, angle = 0),
                       label.theme = element_text(size = 8, angle = 0))) +
    theme_map() +
  facet_wrap(~model, dir = "h", scales = "free_y") +
  panel_border()
```


### Over Time

We plot a line for each year, with a detrended loess line in a darker color.

```{r try a facet}
bottom.row <-  ggplot(data = time.all.models, aes(x = monthLabel, y = pred.mean)) +
    geom_line(aes(group = year), color = "gray70")+
    geom_smooth(aes(x = cal.month, y = pred.mean), se = F, color = "black") +
    theme_minimal() +
    ylab("Mean Probability") +
    xlab("") +
    scale_x_discrete(breaks = c("Jan","Mar","May","Jul","Sep","Nov")) +
    theme(
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      axis.text.x = element_text(size = 10, angle = 90, 
                                 vjust=0.2, hjust = 0),
      panel.background = element_blank(),
      strip.background = element_blank(),
      strip.text.x = element_blank(),
      axis.line = element_line(),
      panel.grid = element_blank(),
      axis.line.y = element_line(),
      axis.ticks.length = unit(0.5, "lines"),
      axis.ticks = element_line()
    ) +
  facet_wrap(~model) +
  panel_border()
```

## Combine into one lattice plot

Use cowplot. Note sometimes you need to make sure any pdf devices are off by running `dev.off()` a whole bunch of times first.

```{r}
pdf(file = "figures/manuscript/Figure4.pdf", width = 6.69, height = 4.75, family = "sans")

ggdraw() +
  draw_plot(top.row, x = 0, y = 0.5, width = 1, height = 0.5) +
  draw_plot(bottom.row, x = 0.02, y = 0, width = 0.98, height = 0.5) +
  draw_plot_label(label = letters[1:6],
                  x = c(0.1,0.4,0.72,0.1,0.4,0.72),
                  y = c(1,1,1,0.55,0.55,0.55), 
                  size = 12)

dev.off()
```

**Caption:** Model variance through space (a,b,c) and time(d,e,f). Plots of variance municipality-level predictions using the single process model (a), dual process model with low reservoir richness (b), and dual process model with high reservoir richness (c). Mean probability across the entire study area of Brazil for the single process model (e), dual process model with low reservoir richness (e), and dual process model with high reservoir richness (f). Gray lines represent an individual year of data with overall mean in black.


# Supp. Figure X: Gif of maps

A seperate GIF for each model. Or maybe one GIF showing all three models at once, which would be more helpful but could be a huge file. 
