---
title: "Manuscript Figures"
author: ""
date: "January 23, 2018"
output: html_document
---

# Packages and Set-Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

#fig 3
library(sf)
library(raster)
library(viridis)
library(cowplot)
library(ggthemes)
```

This document creates the figures for:

*Insert manuscript citation here*

I'm currently loading data individually for each figure, then if you want to remake one figure you don't have to rerun the whole document, just that section.

# Figure 1: ROC Plots

# Figure 2: Variable Importance

Dodged barplot of variable importance

# Figure 3: Lattice Plot 

Figure showing results collapsed over space and time in a lattice. Maybe want to focus on showing variation via sd?. In order for these to play nicely with cowplot, I need to create each facet individually.

## Load Data
```{r}
# spatial data
braz.map <- st_read("../data_clean", "BRAZpolygons")
braz.map <- rename(braz.map, muni.no = muni_no)

#get brazil outline
braz.outline <- getData('GADM', country ='BRA', level = 0)
braz.outline <- st_as_sf(braz.outline)
braz.outline <- dplyr::select(braz.outline , id = OBJECTID, ID_0, ISO)

# prediction data
pred.one.model <- readRDS("../data_out/MS_results/OneModel/wholePredictions.rds")
pred.highNHP <- readRDS("../data_out/MS_results/HighModel/wholePredictions.rds")
pred.lowNHP <- readRDS("../data_out/MS_results/LowModel/wholePredictions.rds")

```

## Calculate Values

### Over Space

I'm calculating the variation per municipality to get a measure of "flickering".

```{r}
# calculate variation per municipality over whole time period
var.one.model <- as.data.frame(pred.one.model) %>%
  group_by(muni.no) %>%
  summarise(pred.var = var(prediction, na.rm = T)) %>%
  mutate(model = "one.model")

var.highNHP <- as.data.frame(pred.highNHP) %>%
  group_by(muni.no) %>%
  summarise(pred.var = var(prediction, na.rm = T)) %>%
  mutate(model = "highNHP")

var.lowNHP <- as.data.frame(pred.lowNHP) %>%
  group_by(muni.no) %>%
  summarise(pred.var = var(prediction, na.rm = T)) %>%
  mutate(model = "lowNHP")

var.all.models <- rbind(var.one.model, var.highNHP, var.lowNHP) %>%
  left_join(braz.map, by = "muni.no")

rm(var.one.model, var.highNHP, var.lowNHP)

var.all.models$model <- factor(var.all.models$model, levels = c("one.model", "lowNHP", "highNHP"))
```


### Over Time

This looks at a time series, averaged over all municipalities by month.

```{r}
time.one.model <- as.data.frame(pred.one.model) %>%
  group_by(month.no) %>%
  summarise(pred.mean = mean(prediction, na.rm = T)) %>%
    mutate(model = "one.model")

time.lowNHP <- as.data.frame(pred.lowNHP) %>%
  group_by(month.no) %>%
  summarise(pred.mean = mean(prediction, na.rm = T)) %>%
    mutate(model = "lowNHP")

time.highNHP <- as.data.frame(pred.highNHP) %>%
  group_by(month.no) %>%
  summarise(pred.mean = mean(prediction, na.rm = T)) %>%
    mutate(model = "highNHP")

time.all.models <- rbind(time.one.model, time.highNHP, time.lowNHP) 

#create cal.month variable (1 - 12)
time.all.models <- time.all.models %>%
  mutate(cal.month = month.no %% 12)

time.all.models$cal.month[time.all.models$cal.month==0] <- 12

#create year variable
time.all.models <- time.all.models %>%
  mutate(year = ((month.no-1) %/% 12) + 2001) %>%
  mutate(monthLabel = month(cal.month, label = T))

#order models
time.all.models$model <- factor(time.all.models$model, levels = c("one.model", "lowNHP", "highNHP"))

rm(time.one.model, time.highNHP, time.lowNHP)
```

## Plot

### Over Space

```{r map theme}
theme_map <- function(...) {
  theme_minimal() +
  theme(
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "left",
    panel.grid = element_blank(),
    plot.background = element_rect(fill = NA, color = NA), 
    panel.background = element_rect(fill = NA, color = NA), 
    legend.background = element_rect(fill = NA, color = NA),
    strip.background = element_blank(),
    strip.text = element_blank(),
    ...
  )
}
```


```{r space plot function}
spacePlot <- function(data.subset, withGuide=T){
  p <- ggplot() +
    geom_sf(data = braz.outline, color = "gray25", size = 0.1) +
    geom_sf(data = data.subset, aes(fill = pred.var), color = NA) +
    coord_sf(datum = NA)+
    scale_fill_viridis(option = "magma", direction = -1, discrete = F,
                     na.value = "gray70", 
                     guide = guide_colorbar(
                       title = "Variation",
                       direction = "vertical",
                       barheight = unit(35, units = "mm"),
                       barwidth = unit(2, units = "mm"),
                       draw.ulim = F,
                       title.position = 'top',
                       # some shifting around
                       title.hjust = 0.5,
                       label.hjust = 0.5,
                       title.theme = element_text(size = 8, angle = 0),
                       label.theme = element_text(size = 8, angle = 0)),
                     limits = c(0,0.07)) +
    theme_map()
  
  #drop guide if we want
  if(withGuide==F){
    p <- p +
      guides(fill = F)
  }
  
  return(p)
}

```


```{r create individual space plots, eval = F}
space.plot.1 <- spacePlot(data.subset = filter(var.all.models, model == "one.model"), withGuide = F)

space.plot.2 <- spacePlot(data.subset = filter(var.all.models, model == "lowNHP"), withGuide = F)

space.plot.3 <- spacePlot(data.subset = filter(var.all.models, model == "highNHP"))
```


```{r facet}
top.row <- ggplot() +
    geom_sf(data = braz.outline, color = "gray25", size = 0.1) +
    geom_sf(data = var.all.models, aes(fill = pred.var), color = NA) +
    coord_sf(datum = NA)+
    scale_fill_viridis(option = "magma", direction = -1, discrete = F,
                     na.value = "gray70", 
                     guide = guide_colorbar(
                       title = "Variation",
                       direction = "vertical",
                       barheight = unit(35, units = "mm"),
                       barwidth = unit(2, units = "mm"),
                       draw.ulim = F,
                       title.position = 'top',
                       # some shifting around
                       title.hjust = 0.5,
                       label.hjust = 0.5,
                       title.theme = element_text(size = 8, angle = 0),
                       label.theme = element_text(size = 8, angle = 0)),
                     limits = c(0,0.07)) +
    theme_map() +
  facet_wrap(~model, dir = "h")
```


### Over Time

We plot a line for each year, with a detrended loess line in a darker color.

```{r}
timePlot <- function(data.subset, yLabel=T){
  p <- ggplot(data = data.subset, aes(x = monthLabel, y = pred.mean)) +
    geom_line(aes(group = year), color = "gray70")+
    geom_smooth(aes(x = cal.month, y = pred.mean), se = F, color = "black") +
    theme_base() +
    ylab("Mean Prediction") +
    xlab("") +
    ylim(0,0.18) +
    scale_x_discrete(breaks = c("Jan","Mar","May","Jul","Sep","Nov")) +
    theme(
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      axis.text.x = element_text(size = 10, angle = 90, vjust=-0.1),
      panel.background = element_blank()
    )
  
  if(yLabel==F){
    p <- p +
      theme(
        axis.title.y = element_blank()
      )
  }
  return(p)
  
}
```


```{r}
time.plot.1 <- timePlot(data.subset = filter(time.all.models, model == "one.model"))
time.plot.2 <- timePlot(data.subset = filter(time.all.models, model == "lowNHP"), yLabel = F)
time.plot.3 <-  timePlot(data.subset = filter(time.all.models, model == "highNHP"), yLabel = F)
```


```{r try a facet}
bottom.row <-  ggplot(data = time.all.models, aes(x = monthLabel, y = pred.mean)) +
    geom_line(aes(group = year), color = "gray70")+
    geom_smooth(aes(x = cal.month, y = pred.mean), se = F, color = "black") +
    theme_base() +
    ylab("Mean Prediction") +
    xlab("") +
    ylim(0,0.18) +
    scale_x_discrete(breaks = c("Jan","Mar","May","Jul","Sep","Nov")) +
    theme(
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      axis.text.x = element_text(size = 10, angle = 90, 
                                 vjust=0.2, hjust = 0),
      panel.background = element_blank(),
      strip.background = element_blank(),
      strip.text.x = element_blank(),
      axis.line = element_line(),
      panel.grid = element_blank(),
      axis.line.y = element_line()
    ) +
  facet_wrap(~model)
```

```{r}
time.plot <- ggplot(data = time.all.models, aes(x=monthLabel, y = pred.mean))+
  geom_line(aes(group = year), color = "gray70")+
  geom_smooth(se = F, color = "black")+
  facet_wrap(~model) +
  theme_bw()

time.plot
```

## Combine into one lattice plot

Use cowplot:

```{r}
pdf(file = "figures/manuscript/Figure3.pdf", width = 6.69, height = 5, family = "sans")

ggdraw()+
  draw_plot(space.plot.1, x = 0, y = 0.5, width = 0.29, height = 0.5) +
  draw_plot(space.plot.2, x = 0.3, y = 0.5, width = 0.29, height = 0.5) +
  draw_plot(space.plot.3, x = 0.59, y = 0.5, width = 0.41, height = 0.5) +
  draw_plot(time.plot.1, x = 0, y = 0, width = 0.32, height = 0.5) +
  draw_plot(time.plot.2, x = 0.33, y = 0, width = 0.32, height = 0.5) +
  draw_plot(time.plot.3, x = 0.66, y = 0, width = 0.32, height = 0.5) +
  draw_plot_label(label = letters[1:6],
                  x = c(0,0.3,0.59,0,0.3,0.59),
                  y = c(1,1,1,0.5,0.5,0.5))

dev.off()
```

Combining already faceted:

```{r}
pdf(file = "figures/manuscript/Figure3.pdf", width = 6.69, height = 5, family = "sans")

ggdraw() +
  draw_plot(top.row, x = 0, y = 0.5, width = 1, height = 0.5) +
  draw_plot(bottom.row, x = 0, y = 0, width = 1, height = 0.5) +
  draw_plot_label(label = letters[1:6],
                  x = c(0,0.25,0.55,0,0.4,0.75),
                  y = c(1,1,1,0.5,0.5,0.5))

dev.off()
```


#Figure 4: Maps

Choose maps at certain time points to represent spatial patterns.

Don't know which time points as of yet. This will pair with the supplementary gif.

# Supp. Figure X: Gif of maps

A seperate GIF for each model? Otherwise it a
