---
title: "Manuscript Figures"
author: ""
date: "January 23, 2018"
output: html_document
---

# Packages and Set-Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(reshape2)
library(sf)
library(raster)
library(viridis)
library(RColorBrewer)
library(cowplot)
library(ggthemes)
library(lubridate)
library(grid)
library(scales)
library(ggsn)
```

Color scheme:
```{r}
singleCol <- alpha("#1b9e77", 0.8)
lrrCol <- alpha("#d95f02",0.8)
hrrCol <- alpha("#7570b3", 0.8)
```

This document creates the figures for:

Kaul, RajReni B., Michelle V. Evans, Courtney C. Murdock, John M. Drake. 2018. Spatio-temporal spillover risk of yellow fever in brazil. *BMC Medicine*.

Data is loaded individually for each figure (except coord plot figure), but packages and set up must be run for everything.

Figure sizing (max file size 10 MB):

Half page width = 85mm (3.3 inch)
Full page width = 170 mm (6.69 inch)
max height = 225 mm (8.85 inch)

# NHP Split

Plots the distribution of NHP richness and the spatial split.

## Load Data

```{r}
braz.outline <- st_read("../data_clean/shapefiles", "brazilOutline")
braz.outline <- dplyr::select(braz.outline, ADMIN)

braz.map <- st_read("../data_clean/shapefiles", "BRAZNoIslands")
braz.map <- rename(braz.map, muni.no = muni_no)

#primateRichness by municipality
primRichness <- readRDS("../data_clean/environmental/primRichness.rds")

braz.map <- left_join(braz.map, primRichness, by = "muni.no")

# NHP split index
nhp.split <- readRDS("../data_clean/environmental/twoModelSplit.rds")
nhp.split.map <- left_join(braz.map, nhp.split, by = "muni.no") %>%
  group_by(above5split) %>%
  summarize() %>%
  #gets out any weird lines between polygons
  st_buffer(0.01)

#create dataframe that keeps split and raw NHP richness
nhpColor <- left_join(primRichness, nhp.split, by = "muni.no")
```

## Plot

### Map

```{r}
splitMap <- ggplot(data = nhp.split.map) +
  geom_sf(aes(fill = above5split)) + 
  scale_fill_manual(values = c(hrrCol, lrrCol)) +
  theme_minimal() +
  coord_sf(xlim = c(-74, -34.78), ylim = c(-33.8, 5.3), datum = NA) +
  guides(fill = F) +
  theme(
    panel.border = element_rect(fill = NA)
  )
```

### Histogram

Based on split

```{r}
histPlot <- ggplot(data = primRichness) +
  geom_histogram(aes(x=spRich), binwidth = 1, fill = c(rep(lrrCol, 6), rep(hrrCol,  17))) +
  ylab(expression("No. of Municipalities")) +
  xlab("NHP Species Richness") +
  scale_y_continuous(trans = 'log10', expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0))
```

Based on species richness
```{r}
histPlot <- ggplot(data = nhpColor) +
  geom_histogram(aes(x=spRich, fill = above5split), binwidth = 1, alpha = 0.8, position = "identity") +
  ylab(expression("No. of Municipalities")) +
  xlab("NHP Species Richness") +
  scale_y_log10(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c(hrrCol, lrrCol)) +
  guides(fill = F)
```

### Facet Together

```{r}
vp <- viewport(width = 0.45, height = 0.45, x = 0.75, y = 0.72)

pdf(file = "../data_out/MS_results/figures/manuscript/nhpSplitDist.pdf", width = 3.3, height = 3.3, family = "sans") 

print(histPlot)
print(splitMap, vp = vp)

dev.off()
```

# Relative Variable Importance as quantiles

The median, and 95% quantile reported for dAUC scaled to max median dAUC for each model. 

```{r relative variable importance median}
#function to calculate dAUC stats
RelativeImportance=function(x, model="NA"){
  perms <- x[[3]][complete.cases(x[[3]]),-c(1, dim(x[[3]])[2])] # pull out AUC values, but drop full model and all permuted data columns 
  model.AUC <- as.numeric(as.character(x[[2]][1,2])) # pull out AUC for full model
  # calculate delta AUC
  delta.AUC <- model.AUC- perms
    # scale delta AUC between zero and one
    #scale.delta.AUC <- delta.AUC/max(delta.AUC)
  #calculate median, upper and lower quantiles
  Quant.dAUC <- delta.AUC %>% 
    melt(value.name="dAUC") %>% 
    rename(Rep=Var1, Variable=Var2)%>% 
    group_by(Variable) %>%
    summarise(dAUCmean=mean(dAUC),dAUCmedian=median(dAUC),Q975=quantile(dAUC, probs=0.975),Q025=quantile(dAUC, probs=0.025) ) %>%
    ungroup()
  
  scale.dAUC <- Quant.dAUC %>%
      group_by(Variable) %>%
      mutate_all(funs(./max(Quant.dAUC$dAUCmedian))) %>% # divid all values by max median value
      arrange(desc(dAUCmedian))%>%
      ungroup() %>%  
      mutate(Rank=c(11:1), Model=model)
      
  return(scale.dAUC)
}

#variable names for publication
niceNames <- c("popLog10"="Population Density",
               "spRich" = "NHP Richness",
               "RFsqrt" = "Mean Rainfall",
               "tempMean" = "Mean Temperature",
               "NDVIScale" = "Scaled NDVI",
               "tempScale" = "Scaled Mean Temperature",
               "primProp" = "NHP Agriculture Overlap",
               "RFScale" = "Scaled Mean Rainfall",
               "fireDenScale" = "Scaled Fire Density",
               "NDVI" ="NDVI",
               "fireDenSqrt" ="Fire Density")

#load data
fullM <- readRDS("../data_out/MS_results/OneModel/Perm100FullModel500TryCatch.rds") 
lowM <- readRDS("../data_out/MS_results/LowModel/lPerm100Model500TryCatch.rds")
highM <- readRDS('../data_out/MS_results/HighModel/HPerm100Model500TryCatch.rds')

#calculate dAUC stats
hri <- RelativeImportance(highM, "Regional: HRR")
lri <- RelativeImportance(lowM, "Regional: LRR")
fri <- RelativeImportance(fullM, "National")
allRelImportance<- rbind(fri,lri,hri)

#reorder factors for plotting 
allRelImportance <- within(allRelImportance, Variable <- factor(Variable, levels = c( "popLog10","spRich","RFsqrt","tempMean","NDVIScale","tempScale","primProp","RFScale","fireDenSqrt","NDVI","fireDenScale")))
allRelImportance <- within(allRelImportance, Model <- factor(Model, levels = c("National","Regional: LRR","Regional: HRR")))

#   
# b2  <-  ggplot(allRelImportance, aes(x=Variable, y=dAUCmedian, ymin = Q025, ymax = Q975, fill=Model))+
#     geom_hline(yintercept = 0.0, color="grey50", linetype=2) +
#     geom_crossbar(position="dodge")+
#     #scale_y_continuous(limits=c(-.5,2))+
#     #geom_point(y=allRelImportance$dAUCmean)+
#     ylab("Relative Importance") + xlab("variable") +
#     scale_x_discrete("",labels=niceNames) +
#     theme_few()+
#     theme(axis.text.x=element_text(angle=35, hjust=1), legend.position = c(.8,.8) ) +
#     scale_fill_tableau("colorblind10")
  
  #tiff("figures/manuscript/VarImpQuantiles.tiff")
  #plot(b2)
  #dev.off()
  
 # ggplot(allRelImportance, aes(x=Variable, y=dAUCmedian, ymin = Q025, ymax = Q975, fill=Model))+
 #    geom_crossbar()+
 #    facet_grid(Model~.)+
 #    scale_y_continuous(limits=c(-.05,1))+
 #    #geom_point(y=allRelImportance$dAUCmean)+
 #    geom_hline(yintercept = 0.0, color="grey50", linetype=2) +
 #    ylab("Relative Importance") + xlab("variable") +
 #    scale_x_discrete("",labels=niceNames) +
 #    theme_few()+
 #    theme(axis.text.x=element_text(angle=35, hjust=1), legend.position = c(.9,.9) ) +
 #    scale_fill_tableau("colorblind10")
 #  
 #  tiff("figures/manuscript/VarImpQuantilesLattice.tiff")
 #  plot(b2)
 #  dev.off()
  
b3 <-  ggplot(allRelImportance, aes(x=Variable, y=dAUCmedian, fill=Model)) + 
  geom_bar(stat="identity",position="dodge")+
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_discrete("Variable",labels=niceNames) +
  theme(text = element_text(face="bold",size=18))+
  theme(axis.text.x=element_text(angle=45, hjust=1), legend.position = c(.6,.8) ) +
  labs(y="Relative Variable Importance") + 
  scale_fill_manual(values = c(singleCol, lrrCol, hrrCol))

plot(b3)

tiff("../data_out/MS_results/figures/manuscript/VarImpMedian.tiff")
plot(b3)
dev.off()

pdf("../data_out/MS_results/figures/manuscript/VarImpMedian.pdf")
plot(b3)
dev.off()
```


# Coord Plot of Variable Importance Rank Median

Ordered based on medians calculated in variable importance plots. Must run code above for data prep.

```{r Coord Plot}

justNamesStart <- c("Population Density",
               "NHP Richness",
               "Mean Rainfall",
               "Mean Temperature",
                "Scaled NDVI",
               "Scaled Mean\nTemperature",
               "NHP Agriculture\nOverlap",
               "Scaled Mean\nRainfall",
                "Fire\nDensity",
               "NDVI",
               "Scaled Fire Density") 

justNamesEnd <- c(
               "NHP Richness",
               "Population Density",
               "Mean Temperature",
               "NHP Agriculture\nOverlap",
               "Scaled NDVI",
               "Mean Rainfall",
               "Scaled Mean\nTemperature",
               "Fire\nDensity",
               "NDVI",
               "Scaled Mean\nRainfall",
               "Scaled Fire Density") 

y_levels <- levels(allRelImportance$Variable)

p <- ggplot(allRelImportance, aes(x = Model, y = Rank, group = Variable)) +   # group = id is important!
  geom_hline(yintercept = c(1:11), color="grey70", linetype=2, alpha=0.5, size=0.25) +
  geom_path(aes(color = Variable, alpha=.95),lineend = 'round', linejoin = 'round', size=3) +
  scale_y_discrete(limits = levels(allRelImportance$Variable), labels=c(11:1)) + #ylim/lab details
  ylab("Ranked Variable Importance") +
  xlab("Model")+
  theme(text = element_text(face="bold",size=18))+
  #theme_few() +
  scale_x_discrete(expand = c(0, 1)) +
  annotate("text", x = c(0.5), y=c(11:1), label = justNamesStart) + 
  annotate("text", x = c(3.5), y=c(11:1), label = justNamesEnd) + 
  theme(legend.position = "none") 

tiff("../data_out/MS_results/figures/manuscript/RankMedian.tiff")
plot(p)
dev.off()

pdf("../data_out/MS_results/figures/manuscript/RankMedian.pdf")
plot(p)
dev.off()
```

#Maps of Calendar Months

This map is a 3 x 2 lattice, showing three time points in 2008 across the national and regional model. These values are represented by percentiles of each individual model.

## Load Data

```{r}
braz.map <- st_read("../data_clean/shapefiles", "BRAZNoIslands")
braz.map <- rename(braz.map, muni.no = muni_no)
braz.map <- st_transform(braz.map, 29101)
# NHP split index
nhp.split <- readRDS("../data_clean/environmental/twoModelSplit.rds")
nhp.split.map <- left_join(braz.map, nhp.split, by = "muni.no") %>%
  group_by(above5split) %>%
  summarize() %>%
  #gets out any weird lines between polygons
  st_buffer(1000)

#get other neighboring countries
s.america <- st_read("../data_clean/shapefiles", "SAmericaCont")

# reproject to SAD69 Brazil Polyconic
s.america <- st_transform(s.america, 29101)

# prediction data
pred.one.model <- readRDS("../data_out/MS_results_revisions/Infection/OneModel/wholePredictions.rds")
pred.highNHP <- readRDS("../data_out/MS_results_revisions/Infection/HighModel/wholePredictions.rds")
pred.lowNHP <- readRDS("../data_out/MS_results_revisions/Infection/LowModel/wholePredictions.rds")
```

## Calculate Data for Plots

Rescale raw outputs into percentiles and combine into one dataframe

```{r}
pred.lowNHP <- pred.lowNHP %>%
  mutate(percentile = cut(prediction, 
                          quantile(prediction, probs = seq(0,1, by = 0.0001)), 
                          include.lowest = T,
                          labels = F)) %>%
  mutate(model = "two.model")

pred.highNHP <- pred.highNHP %>%
  mutate(percentile = cut(prediction, 
                          quantile(prediction, probs = seq(0,1, by = 0.0001)), 
                          include.lowest = T,
                          labels = F)) %>%
  mutate(model = "two.model")

pred.one.model <- pred.one.model %>%
  mutate(percentile = cut(prediction, 
                          quantile(prediction, probs = seq(0,1, by = 0.0001)), 
                          include.lowest = T,
                          labels = F)) %>%
  mutate(model = "one.model")

pred.all <- rbind(pred.lowNHP, pred.highNHP, pred.one.model) %>%
  mutate(cal.month = month.no %% 12)

pred.all$cal.month[pred.all$cal.month==0] <- 12

pred.all.2008 <- pred.all %>%
  mutate(year = case_when(
    cal.month != 12 ~ 2001 + month.no %/% 12,
    cal.month == 12 ~ 2000 + month.no %/% 12)) %>%
  filter(year==2008)

pred.map <- left_join(pred.all.2008, braz.map, by = "muni.no")
```

## Plot

```{r static map theme}
my_theme_map <- function(...) {
  theme_bw() +
  theme(
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "right",
    plot.background = element_rect(fill = NA, color = NA), 
    panel.background = element_rect(fill = NA, color = NA), 
    panel.border = element_rect(fill = NA, color = "black"),
    legend.background = element_rect(fill = NA, color = NA),
    strip.background = element_rect(fill = NA, color = NA),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.spacing = unit(10, "pt"),
    ...
  )
}
```

Subset out months.

```{r}
pred.map.subset <- pred.map %>%
  filter(cal.month %in% c(1,6,9)) %>%
  mutate(monthLabel = factor(month.abb[cal.month], levels = c("Jan", "Jun", "Sep")))
```

Plot.

```{r}

bb <- extent(nhp.split.map)
bb2 <- data.frame(long = unlist(bb[c(1, 2)]), lat = unlist(bb[c(3,4)]))

tiff(file = "../data_out/MS_results_revisions/Infection/figures/manuscript/PercentileMap.tiff", width = 170, height = 110, units = "mm", res = 250, family = "sans", compression = "lzw") #to get less than 10 MB

ggplot() +
  geom_sf(data = s.america, color = "black", fill = "gray80", size = 0.1) +
  geom_sf(data = pred.map.subset, aes(fill = percentile/100), color = NA, size = 0) +
  scale_fill_distiller(palette = "OrRd", direction = 1, type = "seq",
                     guide = guide_colorbar(
                       title = "Relative Risk of\nYF Spillover (%)",
                       direction = "vertical",
                       title.position = 'top',
                       title.theme = element_text(size = 8, angle = 0),
                       label.theme = element_text(size = 8, angle = 0))) +
  facet_grid(model ~ monthLabel, 
             labeller = labeller(
               model = c("one.model" = 'National Model', "two.model" = 'Regional Model')
             )) +
  geom_sf(data = nhp.split.map, fill = NA, color = "black", size=0.1) +
  coord_sf(xlim = c(2788000, 7127000), ylim = c(6256000, 10593000), datum = NA) +
  my_theme_map() +
  ggsn::scalebar(data = bb2, dist = 400, dd2km = F,
                 st.size = 1, height= 0.03, 
                  location = "bottomright"
           ) +
  north(bb2, symbol = 14, scale = 0.1, 
        anchor = c(x = 6950000, y = 7000000))

dev.off()

```

# Lattice Plot 

Figure showing results collapsed over space and time in a lattice. 

## Load Data
```{r}
# spatial data
braz.map <- st_read("../data_clean/shapefiles", "BRAZNoIslands")
braz.map <- rename(braz.map, muni.no = muni_no)
# reproject to SAD69 Brazil Polyconic
braz.map <- st_transform(braz.map, 29101)

# NHP split index
nhp.split <- readRDS("../data_clean/environmental/twoModelSplit.rds")
nhp.split.map <- left_join(braz.map, nhp.split, by = "muni.no") %>%
  group_by(above5split) %>%
  summarize() %>%
  #gets out any weird lines between polygons
  st_buffer(1000)

#get other neighboring countries
s.america <- st_read("../data_clean/shapefiles", "SAmericaCont")
s.america <- st_transform(s.america, 29101)


# prediction data
pred.one.model <- readRDS("../data_out/MS_results_revisions/Infection/OneModel/wholePredictions.rds")
pred.highNHP <- readRDS("../data_out/MS_results_revisions/Infection/HighModel/wholePredictions.rds")
pred.lowNHP <- readRDS("../data_out/MS_results_revisions/Infection/LowModel/wholePredictions.rds")

```

## Calculate Values

### Over Space

This calculates the variance per municipality to get a measure of "flickering". The variance is calculated on the raw values.

```{r}
# calculate variance per municipality over whole time period
var.one.model <- as.data.frame(pred.one.model) %>%
  group_by(muni.no) %>%
  summarise(pred.var = var(prediction)) %>%
  mutate(model = "one.model")

var.highNHP <- as.data.frame(pred.highNHP) %>%
  group_by(muni.no) %>%
  summarise(pred.var = var(prediction)) %>%
  mutate(model = "highNHP")

var.lowNHP <- as.data.frame(pred.lowNHP) %>%
  group_by(muni.no) %>%
  summarise(pred.var = var(prediction)) %>%
  mutate(model = "lowNHP")

var.all.models <- rbind(var.one.model, var.highNHP, var.lowNHP) %>%
  left_join(braz.map, by = "muni.no")

rm(var.one.model, var.highNHP, var.lowNHP)

var.all.models$model <- factor(var.all.models$model, levels = c("one.model", "lowNHP", "highNHP"))
```


### Over Time

This looks at a time series, averaged over all municipalities by month. This is also plotting the raw outputs.

```{r}
time.one.model <- as.data.frame(pred.one.model) %>%
  group_by(month.no) %>%
  summarise(pred.mean = mean(prediction, na.rm = T), 
            sumCases = sum(case)) %>%
  mutate(model = "one.model")

time.lowNHP <- as.data.frame(pred.lowNHP) %>%
  group_by(month.no) %>%
  summarise(pred.mean = mean(prediction, na.rm = T),
            sumCases = sum(case)) %>%
  mutate(model = "lowNHP")

time.highNHP <- as.data.frame(pred.highNHP) %>%
  group_by(month.no) %>%
  summarise(pred.mean = mean(prediction, na.rm = T),
            sumCases = sum(case)) %>%
  mutate(model = "highNHP")

time.all.models <- rbind(time.one.model, time.highNHP, time.lowNHP) 

#create cal.month variable (1 - 12)
time.all.models <- time.all.models %>%
  mutate(cal.month = month.no %% 12)

time.all.models$cal.month[time.all.models$cal.month==0] <- 12

#create year variable
time.all.models <- time.all.models %>%
  mutate(year = ((month.no-1) %/% 12) + 2001) %>%
  mutate(monthLabel = lubridate::month(cal.month, label = T))

#order models
time.all.models$model <- factor(time.all.models$model, levels = c("one.model", "lowNHP", "highNHP"))

rm(time.one.model, time.highNHP, time.lowNHP)
```

## Plot

### Over Space

```{r map theme}
my_theme_map <- function(...) {
  theme_minimal() +
  theme(
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "left",
    panel.grid = element_blank(),
    plot.background = element_rect(fill = NA, color = NA), 
    panel.background = element_rect(fill = NA, color = NA), 
    panel.border = element_rect(fill = NA, color = "black"),
    legend.background = element_rect(fill = NA, color = NA),
    strip.background = element_blank(),
    strip.text = element_blank(),
    plot.margin = unit(c(0,5,0,5), unit = "pt"),
    legend.title.align = 0.5,
    ...
  )
}
```


```{r}
bb <- extent(nhp.split.map)
bb2 <- data.frame(long = unlist(bb[c(1, 2)]), lat = unlist(bb[c(3,4)]))

top.row <- ggplot() +
    geom_sf(data = s.america, color = "black", fill = "gray80", size = 0.1) +
    geom_sf(data = nhp.split.map, color = "black", size = 0.1, fill = NA) +
    geom_sf(data = var.all.models, aes(fill = pred.var), color = NA) +
    scale_fill_distiller(palette = "Greens", direction = 1, type ="seq",
                     na.value = "gray70", 
                     guide = guide_colorbar(
                       title = "Seasonal Variation\nin Spillover\nIntensity",
                       direction = "vertical",
                       barheight = unit(35, units = "mm"),
                       barwidth = unit(4, units = "mm"),
                       draw.ulim = F,
                       title.position = 'top',
                       # some shifting around
                       title.hjust = 0.5,
                       label.hjust = 0.5,
                       title.theme = element_text(size = 8, angle = 0),
                       label.theme = element_text(size = 8, angle = 0))) +
  my_theme_map() +
  facet_wrap(~model, dir = "h") +
  coord_sf(xlim = c(2788000, 7127000), ylim = c(6256000, 10593000), datum = NA) +
  ggsn::scalebar(data = bb2, dist = 400, dd2km = F,
                 st.size = 1, height= 0.03, 
                  location = "bottomright") +
  north(bb2, symbol = 14, scale = 0.1, 
        anchor = c(x = 6950000, y = 7000000))
```


### Over Time

We plot a line for each year, with a detrended loess line in a darker color.

```{r}
bottom.row <-  ggplot(data = time.all.models, aes(x = monthLabel, y = pred.mean)) +
    geom_line(aes(group = year), color = "gray70")+
    geom_smooth(aes(x = cal.month, y = pred.mean), se = F, color = "black") +
    theme_minimal() +
    ylab("Mean Intensity") +
    xlab("") +
    scale_x_discrete(breaks = c("Jan","Mar","May","Jul","Sep","Nov")) +
    theme(
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      axis.text.x = element_text(size = 10, angle = 90, 
                                 vjust=0.2, hjust = 0),
      panel.background = element_blank(),
      strip.background = element_blank(),
      strip.text.x = element_blank(),
      axis.line = element_line(),
      panel.grid = element_blank(),
      axis.line.y = element_line(),
      axis.ticks.length = unit(0.5, "lines"),
      axis.ticks = element_line()
    ) +
  facet_wrap(~model) +
    geom_rug(aes(x=jitter(as.numeric(monthLabel),0.3), alpha = sumCases^0.5, size = sumCases), inherit.aes = FALSE) +
  scale_alpha(range = c(0,1)) +
  scale_size(range = c(0.5,3)) +
  guides(alpha = F, size = F) +
  panel_border()
```

## Combine into one lattice plot

Use cowplot. Note sometimes you need to make sure any pdf devices are off by running `dev.off()` a whole bunch of times first.

```{r}
#pdf(file = "../data_out/MS_results_revisions/figures/manuscript/LatticePlot.pdf", width = 6.69, height = 4.75, family = "sans")
dev.off()
dev.off()

tiff(file = "../data_out/MS_results_revisions/Infection/figures/manuscript/LatticePlot.tiff", width = 2007, height = 1425, units = "px", res = 300, family = "sans", compression = "lzw")


ggdraw() +
  draw_plot(top.row, x = 0, y = 0.5, width = 1, height = 0.5) +
  draw_plot(bottom.row, x = 0.095, y = 0, width = 0.905, height = 0.5) +
  draw_plot_label(label = letters[1:6],
                  x = rep(c(0.16,0.45,0.71), 2),
                  y = rep(c(0.98,0.53), each = 3), 
                  size = 12)

dev.off()
```

# Supp. Figure: GIF of maps

GIF of model predictions as percentiles. This plots indidivudal png files which are then converted to GIF using ImageMagick.

Load Data

```{r}
braz.map <- st_read("../data_clean/shapefiles", "BRAZNoIslands")
braz.map <- rename(braz.map, muni.no = muni_no)
braz.map <- st_transform(braz.map, 29101)
# NHP split index
nhp.split <- readRDS("../data_clean/environmental/twoModelSplit.rds")
nhp.split.map <- left_join(braz.map, nhp.split, by = "muni.no") %>%
  group_by(above5split) %>%
  summarize() %>%
  #gets out any weird lines between polygons
  st_buffer(1000)

#get other neighboring countries
s.america <- st_read("../data_clean/shapefiles", "SAmericaCont")

# reproject to SAD69 Brazil Polyconic
s.america <- st_transform(s.america, 29101)

# prediction data
pred.one.model <- readRDS("../data_out/MS_results_revisions/Infection/OneModel/wholePredictions.rds")
pred.highNHP <- readRDS("../data_out/MS_results_revisions/Infection/HighModel/wholePredictions.rds")
pred.lowNHP <- readRDS("../data_out/MS_results_revisions/Infection/LowModel/wholePredictions.rds")
```

Calculate Percentiles

```{r}
pred.lowNHP <- pred.lowNHP %>%
  mutate(percentile = cut(prediction, 
                          quantile(prediction, probs = seq(0,1, by = 0.0001)), 
                          include.lowest = T,
                          labels = F)) %>%
  mutate(model = "two.model")

pred.highNHP <- pred.highNHP %>%
  mutate(percentile = cut(prediction, 
                          quantile(prediction, probs = seq(0,1, by = 0.0001)), 
                          include.lowest = T,
                          labels = F)) %>%
  mutate(model = "two.model")

pred.one.model <- pred.one.model %>%
  mutate(percentile = cut(prediction, 
                          quantile(prediction, probs = seq(0,1, by = 0.0001)), 
                          include.lowest = T,
                          labels = F)) %>%
  mutate(model = "one.model")

pred.all <- rbind(pred.lowNHP, pred.highNHP, pred.one.model) %>%
  mutate(cal.month = month.no %% 12)

pred.all$cal.month[pred.all$cal.month==0] <- 12

pred.all <- mutate(pred.all, year = 2001 + month.no %/% 12)

pred.map.all <- left_join(pred.all, braz.map, by = "muni.no")
```

Calculate case centroids

```{r}
case.centroids <- pred.map.all %>%
  filter(case==1)

case.centroids <- st_centroid(st_sf(case.centroids))
```

## Plot

Map theme

```{r}
my_theme_map <- function(...) {
  theme_bw() +
  theme(
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "right",
    plot.background = element_rect(fill = NA, color = NA), 
    panel.background = element_rect(fill = NA, color = NA), 
    panel.border = element_rect(fill = NA, color = "black"),
    legend.background = element_rect(fill = NA, color = NA),
    strip.background = element_rect(fill = NA, color = NA),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.spacing = unit(10, "pt"),
    ...
  )
}
```

```{r} 
for (i in 1:max(pred.map.all$month.no)){
  
  png(file = paste0("../data_out/MS_results_revisions/Infection/figures/manuscript/gifs/month", 
                    formatC(i, width = 3, format = "d", flag = "0"), ".png"), 
      width = 600, height = 330, units = "px", res = 100, bg = 'white')
  #adjust for my bad ability at division which makes december respond strange
  if (i%%12 == 0){
    plot.year <- 2000 + (i%/% 12)
    plot.month <- "December"
  } else{
    plot.year <- 2001 + (i%/% 12)
    plot.month <- month.name[i %% 12]
  }
  
  plotTitle <- paste("Spillover risk in", plot.month, plot.year, sep = " ")
  
gif.plot <- ggplot() +
  geom_sf(data = s.america, color = "gray80", fill = "gray80") +
  geom_sf(data = pred.map.all[pred.map.all$month.no==i,], aes(fill = percentile/100), color = NA, size = 0) +
  scale_fill_distiller(palette = "OrRd", direction = 1, type = "seq", limits = c(0,100),
                     guide = guide_colorbar(
                       title = "Percentile",
                       direction = "vertical",
                       title.position = 'top',
                       title.theme = element_text(size = 8, angle = 0),
                       label.theme = element_text(size = 8, angle = 0))
                     ) +
  facet_grid(. ~ model, 
             labeller = labeller(
               model = c("one.model" = 'National', "two.model" = 'Regional')
             )) +
  geom_sf(data = nhp.split.map, fill = NA, color = "black", size=0.1) +
  geom_sf(data = case.centroids[case.centroids$month.no==i,], size = 1) + 
  coord_sf(xlim = c(2788000, 7127000), ylim = c(6256000, 10593000), datum = NA) +
  my_theme_map() +
  ggtitle(plotTitle)

print(gif.plot)

dev.off()

}
```

Use ImageMagick to create plot from png files

cd into `../data_out/MS_results/figures/poster/gifs`
CMDLINE: `convert -delay 50 *.png \MapPercentile.gif` (for gif)
CMDLINE: `convert -antialias -delay 100 *.png MapPercentile.mp4` (for mp4)