---
title: "Results Exploration"
author: "Michelle Evans"
date: "May 31, 2018"
output: html_document
---

This document explores temporal and spatial patterns in the predicions from the national and regional model.


# Package Loading and Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(broom)
```


# Long Term Trends in Data

Longterm trends in predictions were analyzed using slope-tests for each municipality. The prediction for each municipality is assessed overtime, resulting in positive, negative, or no slope. This is done individually for months to get rid of monthly variation.

## Load Data

```{r}
pred.one.model <- readRDS("../data_out/MS_results_revisions/Infection/OneModel/wholePredictions.rds")
pred.highNHP <- readRDS("../data_out/MS_results_revisions/Infection/HighModel/wholePredictions.rds")
pred.lowNHP <- readRDS("../data_out/MS_results_revisions/Infection/LowModel/wholePredictions.rds")

#combine together
pred.one.model$type <- "National"
pred.highNHP$type <- "Regional"
pred.lowNHP$type <- "Regional"
pred.all <- rbind(pred.one.model, pred.highNHP, pred.lowNHP)


# add in year and cal month
pred.all <- pred.all %>%
  mutate(year = ((month.no-1) %/% 12) + 2001) %>%
  mutate(cal.month = month.no %% 12) %>%
  mutate(cal.month = case_when(
    cal.month==0 ~ 12,
    TRUE ~ cal.month
  ))
  
```

## Get Slope Coefficients

```{r}
models <- pred.all %>%
    group_by(cal.month, muni.no, type) %>%
    do(model = lm(prediction ~ year, data = .))
    
tidy.mod <- tidy(models, model) %>%
    filter(term=="year") %>%
    mutate(slope = case_when(
      p.value < 0.05 ~ estimate,
      p.value >= 0.05 ~ 0
    ))
```

Slope tests across all months (instead of by month)

```{r}
month.models <- pred.all %>%
    group_by(muni.no, type) %>%
    do(model = lm(prediction ~ month.no, data = .))

tidy.month.mod <- tidy(month.models, model) %>%
    filter(term=="month.no") %>%
    mutate(slope = case_when(
      p.value < 0.05 ~ estimate,
      p.value >= 0.05 ~ 0
    ))

tidy.month.mod %>%
  group_by(type) %>%
  summarise(none = sum(slope==0), neg = sum(slope<0), pos = sum(slope>0))
```

## Summary of Slope Over Time

```{r}
slope.summary <- tidy.mod %>%
  group_by(type, cal.month) %>%
  summarise(none = sum(slope==0), neg = sum(slope<0), pos = sum(slope>0))
```

```{r}
ggplot(data = slope.summary, aes(y=pos, x = as.factor(cal.month))) +
  geom_point()+
  facet_wrap(~type)
```

## Across Space

Load shapefile and attach to slope data

```{r}
braz.map <- st_read("../data_clean/shapefiles", "BRAZNoIslands")
braz.map <- rename(braz.map, muni.no = muni_no)

#reproject
braz.map <- st_transform(braz.map, 29101)

# NHP split index
nhp.split <- readRDS("../data_clean/environmental/twoModelSplit.rds")
nhp.split.map <- left_join(braz.map, nhp.split, by = "muni.no") %>%
  group_by(above5split) %>%
  summarize() %>%
  #gets out any weird lines between polygons
  st_buffer(dist = 1000)

#link to slope data
braz.slope <- merge(braz.map, tidy.mod, by = "muni.no") %>%
  mutate(monthLabel = factor(month.abb[cal.month], levels = month.abb))
```

Plot (this will be pretty slow, 10 min).

```{r}
tiff(file = "../data_out/MS_results_revisions/Infection/figures/supplement/trendsAcrossSpace.tiff", width = 2007, height = 1425, units = "px", res = 300, family = "sans", compression = "lzw")

braz.slope %>%
  filter(type=="Regional") %>%
  ggplot(data=.) +
    geom_sf(aes(fill = estimate), color = NA) +
    geom_sf(data = nhp.split.map, fill = NA, color = "black", size = 0.1) +
  facet_wrap(~monthLabel, nrow = 4, ncol = 3) +
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0, limits = c(-0.04, 0.04)) +
  ggtitle("Long-term trends in YF intensity by calendar month") +
  theme_bw() +
  coord_sf(datum = NA) +
  labs(fill = "Yearly change in YF\nspillover intensity")

dev.off()
```

Across all 156 months

```{r}
month.slope <- merge(braz.map, tidy.month.mod, by = "muni.no") 

tiff(file = "../data_out/MS_results_revisions/Infection/figures/supplement/trendsAcrossSpace_156months.tiff", width = 2007, height = 1425, units = "px", res = 300, family = "sans", compression = "lzw")

options(scipen = 999)

month.slope %>%
  filter(type=="Regional") %>%
  ggplot(data=.) +
    geom_sf(aes(fill = estimate), color = NA) +
    geom_sf(data = nhp.split.map, fill = NA, color = "black", size = 0.1) +
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0, limits = c(-0.001, 0.001))+
  ggtitle("Long-term trends in YF spillover from 2001 - 2013") +
  theme_bw() +
  coord_sf(datum = NA) +
  labs(fill = "Monthly change in YF\nspillover intensity")

dev.off()
```


# Monthly & Yearly Peaks

Which month had the highest predicted value? Which year?

What about if we average across months (ie all January's)? What if we average across months to get one value for the year?

Note, I'm just going to do this for the regional model.

## Calculate Highs

```{r}
#which of the 156 is highest for that municipality?
month.peak <- pred.all %>%
  filter(type == "Regional") %>%
  group_by(muni.no) %>%
  arrange(desc(prediction)) %>%
  slice(1) %>%
  ungroup() %>%
  dplyr::select(muni.no, peak.month = month.no)

#which calendar month is highest averaged over all 13 years?
cal.month.peak <- pred.all %>%
  filter(type == "Regional") %>%
  group_by(muni.no, cal.month) %>%
  summarise(mean.pred = mean(prediction)) %>%
  arrange(desc(mean.pred)) %>%
  slice(1) %>%
  ungroup() %>%
  dplyr::select(muni.no, peak.cal.month = cal.month)

#highest year (averaging across all months?)
year.peak <- pred.all %>%
  filter(type == "Regional") %>%
  group_by(muni.no, year) %>%
  summarise(mean.pred = mean(prediction)) %>%
  arrange(desc(mean.pred)) %>%
  slice(1) %>%
  ungroup() %>%
  dplyr::select(muni.no, peak.year = year)
```

## Plot Over Time

```{r}
hist(month.peak$peak.month, breaks = 156)

hist(cal.month.peak$peak.cal.month, breaks =12)

hist(year.peak$peak.year, breaks = 13)
```

2008 was definitely the worst year (and 2001). This matches what we know about the El Nino cycles, and really isn't different than what other studies have found (or even just an analysis of cases). Romano etl a. 2014

## Plot Across Space

Load shapefile on line 75.

```{r}
peak.map <- merge(braz.map, month.peak, by = "muni.no") %>%
  merge(cal.month.peak, by = "muni.no") %>%
  merge(year.peak, by = "muni.no") 
```

Peak month no.

```{r}
ggplot(data = peak.map) +
  geom_sf(aes(fill = peak.month), color = NA)+
  geom_sf(data = nhp.split.map, fill = NA, color = "black", size = 0.1) + 
  ggtitle("Peak month per muni (1-156)") +
  scale_fill_distiller(type="seq", palette = 5, direction = -1)
```

Peak calendar month (Jan - Dec.)

```{r}
ggplot(data = peak.map) +
  geom_sf(aes(fill = as.factor(peak.cal.month)), color = NA) +
  geom_sf(data = nhp.split.map, fill = NA, color = "black", size = 0.1) + 
  ggtitle("Peak calendar month averaged from 2001 - 2013") +
  scale_fill_discrete()
```

Peak year

```{r}
ggplot(data = peak.map) +
  geom_sf(aes(fill = peak.year), color = NA) +
  geom_sf(data = nhp.split.map, fill = NA, color = "black", size = 0.1) + 
  ggtitle("Peak year averaged across all 12 months") +
  scale_fill_distiller(type = "qual", palette = 14, direction = 1)
```


# Deprecated



## Using Mixed-Effect Models

Another way is to use a mixed effect model with a random intercept for each municipality and a random slope for each month.

```{r}
pred.one.model$muni.no <- as.factor(pred.one.model$muni.no)
pred.one.model$cal.month <- as.factor(pred.one.model$cal.month)
pred.one.model$time <- pred.one.model$year-2000

mix.mod <- lmer(prediction~ time + (0 + time|cal.month) + (1|muni.no), data = pred.one.model)

summary(mix.mod)
```

Mixed model finds no change overtime, however there is a lot of variance explained by random effects.

```{r}
ranef.muni <- ranef(mix.mod)$muni.no

ranef.month <- ranef(mix.mod)$cal.month
```

*Using a giant mixed model introduces a lot of craziness, probably easier to just make thousands of regressions*

## Giant Linear Model

```{r}
mod <- lm(prediction ~ time:cal.month + time:muni.no, data = pred.one.model)
```

Nope, this is way too large to run on a normal computer.
