---
title: "Case Sparsity"
author: "RK"
date: "June 17, 2017"
output: html_document
---

# Rundown for Meeting with John on 2017-06-20

## Background

This year, there have been 758 confirmed cases of yellow fever in Brazil, more than ten times the maximum number of cases per year since 2000. Furthermore, the majority of these cases fall outside the previous geographic area of vaccine recommendation (it has since been updated), suggesting that transmission here was relatively unexpected, and perhaps due in part to the existence of a naive, unvaccinated population (although we do not have vaccination data to support this hypothesis).

## Question: What municipalities in Brazil are most at risk of yellow fever outbreaks? 

We'd like to approach this from a risk map or suitability perspective, resulting in a heatmap by month of risk of YF cases.

## Data

**Epidemiological Data**

  - Number of cases by month per municipality (referred to as muni) from 2001-2014
  - Positive municipalites (not case number) during 2017 outbreak at three time points (Jan, March, May)

**Environmental Data**

  - Land Cover by year/1km resolution, aggregated to muni
  - Temperature (min, mean, max) by month/1km resolution, aggregated to muni 
  - NDVI by month/1km resolution, aggregated to muni
  - Precipiation by month/4km, aggregated to muni
  - Tree Cover Loss from 2000-2014 (not yearly) for 30x30m pixels

**Socio-demographic Data**

  - Population by muni/year
  - income by muni/year

## Why this meeting?

We now have all this data, and want some guidance on where to go from here and how to visualize and interpret our data to choose a model approach. We imagine this will depend on some characteristics of our dataset, particularly sparseness.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(dplyr)
```


```{r loading data}
YF <- readRDS("YFlong.rds")
```

#Data run down

  The spatial data is broken down into 5564 county equivalent municipalities. Cases are reported monthly from 2001 to 2014 (12months X 14 yrs= 168 months). This means that there are a total of `r (5564*168) ` observations (5564 X 168). Observations without cases could be considered true absences.    

#Exploring the distribution of Yellow Fever 
 
 We'd like to know how sparse the data is, so I will answer the following:
 
    - number of positive observations 
    - average number of cases per year & total cases (overall and by year)
    - number of unique municipalities with cases
    - number of municipalities with multiple positive months (grouped by months, ie those with 2, those with 3, etc.)
    - number of municipalities that never have a yellow fever case
    - when a municipality is positive, the range, mean, median of the number of cases it has during that month

##Temporal breakdown
 
```{r summary stats}
YF$case[is.na(YF$case)] <- 0 #replace case NA with 0

#total number of observations
YF.obs <- YF %>%
          filter(month.no <169, case>0) %>%
          nrow()

sum.na=function(data) {sum(data, na.rm = TRUE)}
ave.na=function(data) {mean(data, na.rm = TRUE)}
sd.na=function(data) {sd(data, na.rm = TRUE)}

#totals for each month.no 
YF.monthly <- YF %>%
              filter(month.no<169) %>% #remove yearly totals
              group_by(month.no) %>% #colaspse by month
              summarise(N=length(na.omit(case)) ,total=sum.na(case), ave=ave.na(case), sd=sd.na(case))

#total for each month in a yr
YF.seasonal <- YF %>%
              filter(month.no<169) %>% #remove yearly totals
              filter(case>0) %>% #drop zero cases
              group_by(cal.month) %>% #within yr then month
              #group_by(cal.month) %>% #colaspse by month
              summarise(N=length(na.omit(case)) ,total=sum.na(case), ave=ave.na(case), sd=sd.na(case))

#totals by year
YF.yearly <- YF %>%
              filter(month.no<169) %>% #remove yearly totals
              filter(case>0) %>%
              group_by(year) %>% #colaspse by month
              summarise(no.muni=length(na.omit(case)) ,total=sum.na(case))

```

  Plot of the entire timeseries. Months with no cases are not plotted. Of the 168 months, only `r nrow(YF.monthly)` months had any reported cases. A majority of those months only had a single case reported. A total of `r sum(YF.monthly$total)` cases were reported over the 14 year period. 

```{r monthly plots, fig.width=10}
par(mfrow=c(1,2))

plot(YF.monthly$total ~ YF.monthly$month.no, xaxt="n", ylab="total cases", xlab="month", type="b")
axis(1, at=seq(12,14*12, by=24))

hist(YF.monthly$total, ylab="# months with X cases", xlab="number of cases reported" , main="")
```

###By year 

There were no cases reported in 2007, 2011, and 2012. A majority of years had fewer than 10 cases. Years with fewer cases also had fewer infected municipalities. 

```{r yearly plots, fig.width=10, fig.height=10}
par(mfrow=c(2,2))
plot(YF.yearly$total ~ YF.yearly$year, pch=19, ylab="Total cases", xlab="Year" )
points(x=c(2007,2011,2012), y=c(0,0,0), pch=19, col=2)

plot(YF.yearly$no.muni ~YF.yearly$year, pch=19,ylab="Infected muni", xlab="Year")
points(x=c(2007,2011,2012), y=c(0,0,0), pch=19, col=2)

plot(YF.yearly$no.muni~YF.yearly$total, xlab="Total cases", ylab="Infected muni")
```

###By calendar month
We can look at the number of cases by month to get at any seasonal trends. December through March has an increase of cases, but when the number of municipalities infected are also taken into account the ave case per month is still very low. January has the most number of infected municipalities over the 14 year timeseries.

```{r seasonal plots, fig.width=10}
par(mfrow=c(1,2))
library(gplots)
plotCI(x=YF.seasonal$cal.month,y=YF.seasonal$ave,  uiw=YF.seasonal$sd , ylab="ave cases per municipality", xlab="month", gap=0)
axis(3, at=YF.seasonal$cal.month, labels=paste("n=", YF.seasonal$N))

hist(YF.seasonal$total, xlab="total cases", ylab="# months with X cases", main="") #average number of cases in X month over the 14 years 
```

##Spatial Breakdown

```{r spatial summary}
#break down of spatial distribution of cases
YF.space <- YF %>%
            filter(month.no<169) %>% #filter out year totals
            filter(case>0) %>% #only look at rows that have cases
            group_by(muni.no) %>% #group by muni
            summarise(N.months=length(case), total.case=sum(case), ave.case=mean(case), sd.case=sd(case))
#pull out those that were only infected once in space (os)
YF.os <- YF.space[which(YF.space$N.months==1),]
#pull out those that were infected multiple times in space (ms)
YF.ms <- YF.space[which(YF.space$N.months>1),]

```

From the 5564 municipalities, `r nrow(YF.space)` municipalities had cases. Of those, `r nrow(YF.os)` muni were only infected once.

```{r spatial plots, fig.width=5, fig.align="center"}
hist(YF.space$N.months, ylab="# muni", xlab="# months with reported cases", main="", breaks=c(0:10))
```

```{r muni infected, fig.width=10}
par(mfrow=c(1,2))
boxplot(YF.space$total.case ~ YF.space$N.months, xlab="# months with reported cases", ylab="total cases reported") -> tmp
axis(3, at=c(1,2,3,4,5), labels=paste("n=",tmp$n), tick=FALSE)

boxplot(YF.space$ave.case ~ YF.space$N.months, xlab="# months with reported cases", ylab="ave cases reported per month") -> tmp
axis(3, at=c(1,2,3,4,5), labels=paste("n=",tmp$n), tick=FALSE)

```

Still working on consecutive months infected and spatial relationship to other infected muni. 

##Recap

 I think it is safe to conclude that the data is very sparse:
 
    - number of positive observations
        *   Of the ~93K observations, only `r YF.obs` had reported YF cases. 
    - average number of cases per year & total cases (overall and by year)
        * Average `r mean(YF.yearly$total)` cases a year, and a total of `r sum(YF.monthly$total)` over 14 years
    - number of unique municipalities with cases
        * From the 5564 municipalities, `r nrow(YF.space)` municipalities had cases. Of those, `r nrow(YF.os)` muni were only infected once.
    - number of municipalities with multiple positive months (grouped by months, ie those with 2, those with 3, etc.)
        * `r nrow(YF.space) - nrow(YF.os)`
    - number of municipalities that never have a yellow fever case
        * `r 5564-nrow(YF.space)`
    - when a municipality is positive, the range, mean, median of the number of cases it has during that month
        * see above plots